Multiple regression: TRELIS-Q WP5 data

### 2024-02-15 Rmd/bookchapter.TRELIS-Q-WP5.Rmd
Checked interaction effects more closely

```{r TK_gender_trinn}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Trinn),
                 trace.factor = factor(data_clean$Gender),
                 response = data_clean$TK,
                 ylab = "TK",
                 xlab = "Target grade level",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Gender")
```

@@

@@

@@

## Hierarchical regression analysis

@@

```{r level1}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m1 <- lm_robust(scale(TPACK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m1)
```

@@

```{r level2}
# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
m2 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) * (factor(semester) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m2)
```

@@

@@

```{r PCK}
# PCK
m3 <- lm_robust(TPACK ~ factor(Gender) * (factor(Trinn) + factor(semester)) +         # level 1
                        factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
                        PCK, 
                clusters = factor(studiested), data=data_clean)
summary(m3)

fstat <- m3$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

@@

```{r TK}
# TK
m4 <- lm_robust(TPACK ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                        factor(studiested) * (factor(semester) + factor(Gender))     # level 2
                        + PCK + TK, 
                clusters = factor(studiested), data=data_clean)
summary(m4)

>>>>

### TK Level 1

```{r TK_gender_trinn}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$Trinn),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Target grade level")
```

@@

```{r TK_gender_semester}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Year of study")
```

```{r TK_semester_trinn}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Trinn),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Target grade level",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Year of study")
```

### TK Level 2

```{r TK_semester_institute}
interaction.plot(x.factor = factor(data_clean$semester), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

```{r TK_trinn_institute}
interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

```{r TK_gender_institute}
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$studiested), #label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$TK,
                 type = "b",
                 ylab = "TK",
                 xlab = "Gender",
                 col = viridis(5), #c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

### PCK Level 1

```{r PCK_gender_trinn}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$Trinn),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Target grade level")
```

Different signs on slope, a clear indication of interaction effects on PCK from gender and target grade level.

```{r PCK_gender_semester}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Gender",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Year of study")
```

```{r PCK_semester_trinn}
# ref https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/#viridis-color-palettes
# https://www.statology.org/interaction-plot-r/
interaction.plot(x.factor = factor(data_clean$Trinn),
                 trace.factor = factor(data_clean$semester),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Target grade level",
                 col = viridis(2), # c("pink", "blue"), 
                 trace.label = "Year of study")
```

### PCK Level 2

```{r PCK_semester_institute}
interaction.plot(x.factor = factor(data_clean$semester), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

```{r PCK_trinn_institute}
interaction.plot(x.factor = factor(data_clean$Trinn), # label=c("Year 2/3", "Year 5")),
                 trace.factor = factor(data_clean$studiested), # label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Year of study",
                 col = viridis(5), # c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

```{r PCK_gender_institute}
interaction.plot(x.factor = factor(data_clean$Gender),
                 trace.factor = factor(data_clean$studiested), #label=c("institute 1", "institute 2", "institute 3", "institute 4", "institute 5")),
                 response = data_clean$PCK,
                 type = "b",
                 ylab = "PCK",
                 xlab = "Gender",
                 col = viridis(5), #c("red", "green", "orange", "pink", "blue"),
                 trace.label = "Institute")
```

### TPACK Level 1

@@

### TPACK Level 2

@@

## Hierarchical RA for TK

@@

```{r TKlevel1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m1 <- lm_robust(TK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m1)
```

```{r TKlevel1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(TK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m2)
```

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / m1$adj.r.squared
df1 <- length(data_clean$TK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

@@

```{r TKlevel2_main}
# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
m3 <- lm_robust(scale(TK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m3)
```

```{r}
# main
fstat <- m3$adj.r.squared / m1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m3$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r TKlevel2_interaction}
# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(TK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)
```

```{r}
# interaction
fstat <- m4$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

## Hierarchical RA for PCK

```{r PCKlevel1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m1 <- lm_robust(PCK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m1)
```

```{r PCKlevel1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(PCK, center=T, scale=T) ~ factor(semester) * (factor(Gender) + factor(Trinn)), 
                clusters = factor(studiested), data=data_clean)
summary(m2)
```

```{r}
# source: https://www.statology.org/how-to-calculate-the-p-value-of-an-f-statistic-in-r/
# source: p. 257 i Howitt & Cramer, 2014
fstat <- m2$adj.r.squared / m1$adj.r.squared
df1 <- length(data_clean$PCK) - 1
df2 <- df1
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r PCKlevel2_main}
# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
m3 <- lm_robust(scale(PCK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) + (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m3)
```

```{r}
# main
fstat <- m3$adj.r.squared / m1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m3$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r PCKlevel2_interaction}
# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(PCK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) * (factor(semester) + factor(Trinn) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)
```

```{r}
fstat <- m4$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

## Hierarchical RA for TPACK

```{r level1_main}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m1 <- lm_robust(TPACK ~ factor(Gender) + (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m1)
```

```{r level1_interaction}
# kovariater; we use scale to standardize when using interaction effects
# sjekk: er det riktig å skalere?
m2 <- lm_robust(scale(TPACK, center=T, scale=T) ~ factor(Gender) * (factor(Trinn) + factor(semester)), 
                clusters = factor(studiested), data=data_clean)
summary(m2)
```

@@

```{r level2_main}
# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
m3 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) + (factor(semester) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m3)

#main
fstat <- m3$adj.r.squared / m1$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m3$adj.r.squared / m2$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

```{r level2_interaction}
# studiested; HVL velges som referansekategori siden den har lavest TPACK-verdi av alle
m4 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                               factor(studiested) * (factor(semester) + factor(Gender)),    # level 2
                clusters = factor(studiested), data=data_clean)
summary(m4)

@@

```{r PCK}
# PCK
m5 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +         # level 1
                        factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
                        scale(PCK), 
                clusters = factor(studiested), data=data_clean)
summary(m5)

# main
fstat <- m5$adj.r.squared / m3$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)

# interaction
fstat <- m5$adj.r.squared / m4$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```

@@

```{r TK}
# TK
m6 <- lm_robust(scale(TPACK) ~ factor(Gender) * (factor(Trinn) + factor(semester)) +        # level 1
                        factor(studiested) * (factor(semester) + factor(Gender)) +    # level 2
                        scale(PCK) + scale(TK), 
                clusters = factor(studiested), data=data_clean)
summary(m6)

fstat <- m6$adj.r.squared / m5$adj.r.squared
pf(fstat, df1, df2, lower.tail = FALSE)
```
